### 2022-11-25 paper2.tex
paper 2

\section{Introduction}
Replication studies are conducted in order to investigate whether an original
finding can be confirmed in an independent study. Although replication has long
been a central part of the scientific method in many fields, the so-called
replication crisis \citep{Ioannidis2005, Begley2015} has led to increased
interest in replication over the last decade. These developments eventually
culminated in large-scale replication projects that were conducted in various
fields \citep{Errington2014, Klein2014, OSC2015, Ebersole2016, Camerer2016,
  Camerer2018, Cova2018, Klein2018}.

% Declaring a replication as successful is, however, not a
% straightforward task, and currently used approaches include
% \emph{statistical} significance of both the original and replication
% study,
%  compatibility of their effect estimates, and meta-analysis of
% the effect estimates. \emph{Many of the replication projects listed above also
%   report the relative effect size, the ratio of the
%   replication to the original effect estimate. For example, in \citet{Camerer2018}
%   the replication effect estimates were only half as large as the original ones
%   on average and even smaller in \citet{OSC2015}. This gives
%   clear evidence of a systematic bias of the original studies and
%   strongly suggests that the original and replication study should not
%   be treated as exchangeable. However, all the approaches mentioned
%   above will give the same results if the order of studies would be reversed. }

% In order to address \emph{this problem}, a new method has recently been
% proposed in \citet{held2020}. The approach combines the analysis of
% credibility \citep{matthews:2001,matthews:2001b} with a prior-data
% conflict assessment \citep{box:1980}. Replication success is declared if
% the replication study is in conflict with a sceptical prior that would
% make the original study non-significant. \emph{This approach penalizes
%   small relative effect sizes as we will see in more detail in the following.}

% To introduce some notation, let $z_o = \hat \theta_o/\sigma_o$ and
% $z_r = \hat \theta_r/\sigma_r$ denote the $z$-statistic of the
% original and replication study, respectively. Here $\hat \theta_o$ and
% $\hat \theta_r$ are the corresponding effect estimates (assumed to be
% normally distributed) of the unknown
% effect $\theta$ with standard errors $\sigma_o$ and $\sigma_r$, respectively.
% The corresponding one-sided
% $p$-values are denoted by $p_o=1-\Phi(z_o)$ and $p_r=1-\Phi(z_r)$,
% respectively, where $\Phi(\cdot)$ denotes the standard normal cumulative
% distribution function.
% Let $c = \sigma_o^2/\sigma_r^2$ denote the variance
% ratio of the squared standard errors of the original and replication
% effect estimates. The squared standard errors are usually inversely
% proportional to the sample size of each study, \ie
% $\sigma_o^2 = \kappa^2/n_o$ and $\sigma_r^2 = \kappa^2/n_r$ for some
% unit variance $\kappa^2$.  The variance ratio $c$ can then be
% identified as the relative sample size $c=n_r/n_o$. The relative effect size
% \begin{equation}\label{eq:d}
%   d = \frac{\hat \theta_r}{\hat \theta_o} = \frac{1}{\sqrt{c}} \frac{z_r}{z_o}
% \end{equation}
% quantifies the size of the replication effect estimate $\hat \theta_r$ relative
% to the original effect estimate $\hat \theta_o$. The corresponding shrinkage
% of the replication effect estimate will be denoted as $s=1-d$.

% Suppose the original study achieved statistical significance at one-sided
% level $\alpha$, so $p_o \leq \alpha$.  The standard approach to assess
% replication success is based on significance of the replication effect
% estimate at the same level $\alpha$, \ie the replication is considered successful if also $p_r \leq \alpha$.
% This approach is known in drug development as the two-trials rule \citep{senn:2007}, usually
% conducted at $\alpha=0.025$.
% Let $z_\alpha = \Phi^{-1}(1-{\alpha})>0$ denote the $z$-value corresponding
% to the level $\alpha$, then significance of the replication
% \emph{study is achieved if $z_r \geq z_\alpha$, which}
% is equivalent to the \emph{condition}
% \begin{equation}\label{eq:dSig}
%    d \geq \frac{z_\alpha}{z_o  \, \sqrt{c}}
% \end{equation}
% on the relative effect size \eqref{eq:d}.
% The right hand-side goes to zero for increasing $c$, so if the
% relative sample size $c$ is large enough, significance of the
% replication study can be achieved with any arbitrarily small (but
% positive) relative effect size $d$. \emph{However, declaring replication
% success when there is substantial shrinkage is contrary to common
% sense, as the replication effect estimate may not reflect an effect
% size of the same practical relevance as the original one, despite its
% statistical significance.}

% \emph{In this paper we first review the \citet{held2020} approach for
%   the assessment of replication success, followed by showing how it
%   relates to the relative effect size (Section \ref{sec:res}). This perspective
%   is used in Section \ref{sec:goldenthresh} and \ref{sec:recalib}
%   to propose  a recalibration of the
%   method, the \textit{golden level}, which leads to a more appropriate
%   criterion for replication success compared to the two-trials rule
%   (Section \ref{sec:2TR}). In Section \ref{sec:ER} we study
%   power and Type-I error rates of the proposed method and compare it to the
%   two-trials rule. The recalibrated method ensures that conditional power can
%   take any desired value if the original study has been significant and the replication
%   sample size is large enough
% (Section \ref{sec:powerrep}), controls the overall Type-I error if the replication
% sample size is not smaller than the original one (Section \ref{sec:T1E}), and
% offers uniform
% gains in project power compared to the two-trials rule (Section \ref{sec:PP}). Section \ref{sec:application}
% describes an application to data from four replication projects and Section
% \ref{sec:discussion} closes with some discussion.}

% \clearpage
% \section{Replication success}\label{sec:RS}

% \begin{center}
% \begin{figure}[!h]
% \begin{center}
% \begin{knitrout}
% \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

% {\centering \includegraphics[width=\maxwidth]{figure/fig1-1}

% }

% \end{knitrout}
% \caption{Example of the assessment of replication success. The original study from \citet{Pyc2010} has effect estimate  $\hat \theta_o=
%   0.4$ on Fisher's $z$ scale
%   (95\% CI from $0.05$ to $0.74$) and one-sided $p$-value
%   $p_o = 0.011$.
%   The left part of the figure illustrates the \emph{reverse-Bayes derivation of the sufficiently sceptical prior
%     based on}
%   the original study result and the posterior with lower credible limit fixed at zero. \emph{The comparison} of the sufficiently sceptical prior with the replication study result ($\hat \theta_r = 0.15$, 95\% CI from 0.04 to 0.26, $p_r = 0.004$)
%   in the right part of the figure \emph{is used to assess potential prior-data conflict.}
%   \label{fig:fig1}}
% \end{center}
% \end{figure}
% \end{center}

% Hereinafter we focus on the one-sided assessment of replication
% success to ensure that replication success can only occur if the
% original and replication effect estimates go in the same
% direction. Figure \ref{fig:fig1} illustrates the \citet{held2020} approach
% based on a replication study
% from the \textit{Social Sciences Replication Project} \citep{Camerer2018}:
% the significant original finding by \citet{Pyc2010} at
% one-sided level $\alpha=0.025$ is
% challenged with a sceptical prior, sufficiently concentrated around
% zero to make the original study result no longer convincing \citep{matthews:2001,matthews:2001b}.
% Replication success is then defined as conflict between the
% sceptical prior and the result from the replication study in order to
% \emph{disprove} the sceptic. Conflict is quantified by a prior-predictive
% tail probability $p_{\mbox{\scriptsize Box}}$ \citep{box:1980} where a small value
% $p_{\mbox{\scriptsize Box}} \leq \alpha$ defines replication success.
% In Figure \ref{fig:fig1}  the original finding is only borderline significant, so the
% sufficiently sceptical prior is fairly wide.  Furthermore, there is
% substantial shrinkage
% ($d = 0.15/0.4 =
% 0.38$)
% of the replication effect estimate
% and therefore hardly any conflict with the sufficiently
% sceptical prior (one-sided $p_{\mbox{\scriptsize Box}}=0.31$).
% We are thus not able to declare replication success at
% level $2.5$\%.

% \emph{The actual value of $p_{\mbox{\scriptsize
%       Box}}$ is difficult to interpret as it depends on the level $\alpha$ and does not even exist if
%   the original $p$-value $p_o$ exceeds $\alpha$. However, }
% \citet{held2020} showed
% that if both $\sign(z_o)=\sign(z_r)$ and
% \begin{equation}\label{eq:extrinsic.p}
% \left({z_o^2}/{z_{{\alpha_S}}^2}-1 \right) \left({z_r^2}/{z_{{\alpha_S}}^2}
% - 1 \right) \geq c
% \end{equation}
% hold, replication success at level ${\alpha_S}$ is achieved, where $z_{{\alpha_S}} = \Phi^{-1}(1-{\alpha_S})$.
% The requirement \eqref{eq:extrinsic.p} can be assessed for \emph{any} value of
% \emph{${{\alpha_S}} > \max\{p_o, p_r\}$} and of particular interest is the
% smallest possible value of $\alpha_S$ where \eqref{eq:extrinsic.p} holds,
% the so-called {\it sceptical $p$-value} $p_S$.
% We are thus interested in the value $z_S^2$ that fulfills
% \begin{equation}\label{eq:extrinsic.p2}
% \left({z_o^2}/{z_{S}^2}-1 \right) \left({z_r^2}/{z_{S}^2}
% - 1 \right) = c.
% \end{equation}
% There is a unique solution of \eqref{eq:extrinsic.p2} which defines
% the one-sided {sceptical $p$-value} $p_S=
% 1-\Phi\left({z_S}\right)$ where $z_{S} \coloneqq +
% \sqrt{z_{S}^2}$, provided
% $\sign(z_o)=\sign(z_r)$ holds.  Replication success at level
% $\alpha_S$ is then achieved if $p_S \leq
% \alpha_S$.  In the introductory example based on the original study by
% \citet{Pyc2010}, the sceptical
% $p$-value turns out to be $p_S=0.11$.

% The sceptical
% $p$-value has a number of interesting properties, see \citet[Section
% 3.1]{held2020} for details. In particular, $p_S > \max\{p_o, p_r\}$
% always holds with $p_S \downarrow \max\{p_o, p_r\}$ for $c \downarrow
% 0$. Furthermore, if the $p$-values $p_o$ and
% $p_r$ are fixed, the sceptical $p$-value
% $p_S$ increases with \emph{decreasing} relative effect size
% $d$.  The first property ensures that both the original and the
% replication study have to be sufficiently convincing on their own to
% \emph{achieve} replication success.  The second property guarantees that
% shrinkage of the replication effect estimate is penalized.

% The level for replication success $\alpha_S$ has to be distinguished
% from the significance level $\alpha$ associated with the ordinary
% $p$-value.  \citet{held2020} has used the {\it nominal level} for
% replication success ($\alpha_S=\alpha$) for convenience, but in the
% following we will propose a recalibration of the procedure along with
% a new value for $\alpha_S$, the {\it golden level} (Section~\ref{sec:goldenthresh}).
% \emph{The derivation is based on a property of the required
% relative effect size for replication success, if the relative sample size is very large
% (Section~\ref{sec:res}).}
% In a nutshell, the golden level
% ensures that for original studies which were only borderline
% significant \emph{($p_o=\alpha$)}, replication success is only possible if the replication
% effect estimate is larger than the original one \emph{($d > 1$)}.\\

% \subsection{Relative effect size}\label{sec:res}
% Without loss of generality we \emph{now} assume that $\hat \theta_o > 0$ and that
% $p_o < {\alpha_S}$ has been observed in the original study, otherwise
% it would be impossible to achieve replication success at level
% $\alpha_S$ because $p_S$ is always larger than $p_o$.
% The condition \eqref{eq:extrinsic.p} for replication
% success can then be re-written as
% \begin{equation}\label{eq:eq.z}
%   z_r \geq   z_{{\alpha_S}} \sqrt{1+c/(K-1)} \eqqcolon  \zrmin,
% \end{equation}
% where $K=z_o^2/z_{{\alpha_S}}^2>1$. The right hand-side of
% \eqref{eq:eq.z} is the minimum replication $z$-value $\zrmin$ required
% to achieve replication success.  Note that $\zrmin$ increases with increasing $c$, so increasing the replication
% sample size leads to a more stringent success requirement \emph{for $z_r$ and the corresponding} replication
% $p$-value $p_r$.

% Equation \eqref{eq:eq.z} can be further transformed to a
% condition on the relative effect size \eqref{eq:d}:
% \begin{equation}\label{eq:res}
%   d \geq  \frac{\sqrt{1+c/(K-1)}}{ \sqrt{c K}} \eqqcolon \dmin .
% \end{equation}
% To achieve replication success, the relative effect size
% must be
% at least as large as the right hand-side of \eqref{eq:res}, the
%   minimum relative effect size $\dmin$, a function of $K$ and the
% relative sample size $c$.
% If the relative sample
% size becomes very large, \ie $c \rightarrow \infty$, we have
% $\dmin \downarrow \dinfty$ where
% \begin{equation}\label{eq:smallestpossible}
%   \dinfty = 1/\sqrt{K (K-1)}
% \end{equation}
% \emph{is the {\it limiting relative effect size}.}
% This shows that the minimum relative effect size $\dmin$ in \eqref{eq:res}
% does not go to zero for increasing $c$, so replication success cannot
% be achieved if the relative effect size $d$ is smaller or equal to
% $\dinfty$, no matter how large the replication study is.
% \emph{In contrast}, the corresponding criterion \eqref{eq:dSig} \emph{of the
% two-trials rule}
% can be achieved for any positive relative effect size, regardless of
% how small, provided the replication sample size is sufficiently large.

% \subsection{The golden level}\label{sec:goldenthresh}
% Significance of both the original and the replication study at level
% $\alpha$ is a necessary but not sufficient requirement for replication
% success at the nominal level ($\alpha_S = \alpha$).  The nominal level
% may therefore be too stringent.  It is more reasonable to calibrate
% the procedure in such a way that to establish replication success,
% original and replication study do not both necessarily need to be
% significant at level $\alpha$, provided that the replication effect
% estimate does not shrink compared to the original one.  We therefore
% choose a level $\alpha_S$ such that a borderline
% significant original study ($p_o = \alpha$) cannot lead to replication
% success if there is shrinkage $s > 0$ of the replication effect
% estimate.  Mathematically, this translates to setting $\dinfty = 1$
% and $K = z_\alpha^2/z_{\alpha_S}^2$ in~\eqref{eq:smallestpossible} and
% leads to the quadratic equation $K(K-1) = 1$ with solution
% $K = \varphi$ where
% $\varphi = (\sqrt{5}+1)/2 \approx 1.62$ is
% known as the golden ratio. Solving for $z_{\alpha_S}$ gives
% \emph{$z_{\alpha_S} = z_\alpha / \sqrt{\varphi}$
% and the corresponding  \textit{golden level}
% \begin{eqnarray}\label{eq:golden}
% \alpha_S & = & 1 -  \Phi(z_\alpha / \sqrt{\varphi} )
% \end{eqnarray}
% for replication success. }
% This is our recommended default choice to assess replication success and we will study its properties in the following
%                    in more detail. For $z_\alpha = 1.96$ (one-sided $\alpha = 0.025$),
% the golden level is $\alpha_S =0.062$.
% In the introductory example
% shown in Figure \ref{fig:fig1},
% the sceptical $p$-value is $p_S=0.11 > 0.062$, so
% the replication  of the \citet{Pyc2010} study was not successful.

% \emph{The golden level \eqref{eq:golden} is derived from~\eqref{eq:smallestpossible} with $\dinfty = 1$.
% However, we may also use a different value for the
% limiting relative effect size $\dinfty$, say $\dinfty=0.8$.
% Then replication success is only possible for a borderline significant result ($p_o = \alpha$)
% if there is less than $1-\dinfty$ (20\% for $\dinfty=0.8$) shrinkage of the replication effect estimate.
% This approach is equivalent to a limiting relative effect size of 1 if the original $p$-value
% $p_o$ is equal to a different level $\alpha'$, which can be derived as follows:
% First, solving \eqref{eq:smallestpossible} for $\dinfty > 0$ gives
% $K = z_\alpha^2/z_{\alpha_S}^2 = 1/2+\sqrt{1/4 + 1/\dinfty^2}$. The new level
% $\alpha'$ fulfills  $\varphi=z_{\alpha'}^2/z_{\alpha_S}^2$, so $z_\alpha^2/K=z_{\alpha'}^2/\varphi$
% and therefore
% \begin{equation}\label{eq:alphaPrime}
% \alpha' = 1-\Phi\left(z_{\alpha} \, \sqrt{\varphi / K} \right).
% \end{equation}
% For example, for $\alpha = 0.025$ and $\dinfty = 0.8$ we obtain $\alpha'=0.033$.
% }

% \subsection{Recalibration of the sceptical $p$-value}\label{sec:recalib}

% The condition $p_S \leq \alpha_S$ for replication success at the
% golden level is equivalent to $z_S \geq z_\alpha / \sqrt{\varphi}$,
% \ie $z_S \sqrt{\varphi} \geq z_\alpha$.  In practice it may be
% preferable to recalibrate the sceptical $p$-value
% $p_{S} = 1 - \Phi(z_S)$ to
% $ \tilde{p}_S = 1 - \Phi(z_S \sqrt{\varphi})$, which then needs to be
% compared to $\alpha$ (rather than $\alpha_S$) to assess replication
% success and can thus be interpreted on the same scale as an ordinary
% $p$-value.  For example, the recalibrated sceptical $p$-value for the
% replication of \citet{Pyc2010} turns out to be
% $\tilde{p}_S=0.061$ and does not lead to replication
% success at \emph{any level $\alpha<0.061$, including} the
% standard $0.025$ level.

% \subsection{Comparison with the two-trials rule}\label{sec:2TR}
% A useful benchmark for comparison is the two-trials rule in drug
% development \citep[Section 9.4]{kay:2015}, which requires ``at least
% two adequate and well-controlled studies, each convincing on its own,
% to establish effectiveness'' \citep[p.~3]{FDA1998}.  This is usually
% achieved by independently replicating the result of a first study in a
% second study, both significant at one-sided level
% $\alpha=0.025$.  {It is worth noting that in practice the two trials are often run in parallel} {\citep{senn:2007}}{, so do not exactly resemble the replication setting.}

>>>>

\section{Introduction}
Replication studies are conducted in order to investigate whether an original
finding can be confirmed in an independent study. Although replication has long
been a central part of the scientific method in many fields, the so-called
replication crisis \citep{Ioannidis2005, Begley2015} has led to increased
interest in replication over the last decade. These developments eventually
culminated in large-scale replication projects that were conducted in various
fields \citep{Errington2014, Klein2014, Opensc2015, Ebersole2016, Camerer2016,
  Camerer2018, Cova2018, Klein2018}.

Declaring a replication as successful is, however, not a straightforward task,
and currently used approaches include statistical significance of both the
original and replication study, compatibility of their effect estimates, and
meta-analysis of the effect estimates. Many of the replication projects listed
above also report the relative effect size, the ratio of the replication to the
original effect estimate. For example, in \citet{Camerer2018} the replication
effect estimates were only half as large as the original ones on average and
even smaller in \citet{Opensc2015}. This gives clear evidence of a systematic
bias of the original studies and strongly suggests that the original and
replication study should not be treated as exchangeable. However, all the
approaches mentioned above will give the same results if the order of studies
would be reversed.

In order to address this problem, a new method has recently been proposed in
\citet{Held2020}. The approach combines the analysis of credibility
\citep{Matthews2001a,Matthews2001b} with a prior-data conflict assessment
\citep{Box1980}. Replication success is declared if the replication study is in
conflict with a sceptical prior that would make the original study
non-significant. This approach penalizes small relative effect sizes as we will
see in more detail in the following.

To introduce some notation, let $z_o = \hat \theta_o/\sigma_o$ and
$z_r = \hat \theta_r/\sigma_r$ denote the $z$-statistic of the original and
replication study, respectively. Here $\hat \theta_o$ and $\hat \theta_r$ are
the corresponding effect estimates (assumed to be normally distributed) of the
unknown effect $\theta$ with standard errors $\sigma_o$ and $\sigma_r$,
respectively. The corresponding one-sided $p$-values are denoted by
$p_o=1-\Phi(z_o)$ and $p_r=1-\Phi(z_r)$, respectively, where $\Phi(\cdot)$
denotes the standard normal cumulative distribution function. Let
$c = \sigma_o^2/\sigma_r^2$ denote the variance ratio of the squared standard
errors of the original and replication effect estimates. The squared standard
errors are usually inversely proportional to the sample size of each study, \ie
$\sigma_o^2 = \kappa^2/n_o$ and $\sigma_r^2 = \kappa^2/n_r$ for some unit
variance $\kappa^2$. The variance ratio $c$ can then be identified as the
relative sample size $c=n_r/n_o$. The relative effect size
\begin{equation}\label{eq2:d}
  d = \frac{\hat \theta_r}{\hat \theta_o} = \frac{1}{\sqrt{c}} \frac{z_r}{z_o}
\end{equation}
quantifies the size of the replication effect estimate $\hat \theta_r$ relative
to the original effect estimate $\hat \theta_o$. The corresponding shrinkage of
the replication effect estimate will be denoted as $s=1-d$.

Suppose the original study achieved statistical significance at one-sided level
$\alpha$, so $p_o \leq \alpha$. The standard approach to assess replication
success is based on significance of the replication effect estimate at the same
level $\alpha$, \ie the replication is considered successful if also
$p_r \leq \alpha$. This approach is known in drug development as the two-trials
rule \citep{Senn2008}, usually conducted at $\alpha=0.025$. Let
$z_\alpha = \Phi^{-1}(1-{\alpha})>0$ denote the $z$-value corresponding to the
level $\alpha$, then significance of the replication study is achieved if
$z_r \geq z_\alpha$, which is equivalent to the condition
\begin{equation}\label{eq2:dSig}
   d \geq \frac{z_\alpha}{z_o  \, \sqrt{c}}
\end{equation}
on the relative effect size \eqref{eq2:d}. The right hand-side goes to zero for
increasing $c$, so if the relative sample size $c$ is large enough, significance
of the replication study can be achieved with any arbitrarily small (but
positive) relative effect size $d$. However, declaring replication success when
there is substantial shrinkage is contrary to common sense, as the replication
effect estimate may not reflect an effect size of the same practical relevance
as the original one, despite its statistical significance.

In this paper we first review the \citet{Held2020} approach for the assessment
of replication success, followed by showing how it relates to the relative
effect size (Section \ref{sec2:res}). This perspective is used in Section
\ref{sec2:goldenthresh} and \ref{sec2:recalib} to propose a recalibration of the
method, the \textit{golden level}, which leads to a more appropriate criterion
for replication success compared to the two-trials rule (Section \ref{sec2:2TR}).
In Section \ref{sec2:ER} we study power and Type-I error rates of the proposed
method and compare it to the two-trials rule. The recalibrated method ensures
that conditional power can take any desired value if the original study has been
significant and the replication sample size is large enough (Section
\ref{sec2:powerrep}), controls the overall Type-I error if the replication sample
size is not smaller than the original one (Section \ref{sec2:T1E}), and offers
uniform gains in project power compared to the two-trials rule (Section
\ref{sec2:PP}). Section \ref{sec2:application} describes an application to data
from four replication projects and Section \ref{sec2:discussion} closes with some
discussion.

\section{Replication success}\label{sec2:RS}

\begin{center}
\begin{figure}[!h]
\begin{center}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\maxwidth]{images/paper2/fig1-1}

}

\end{knitrout}
\caption{Example of the assessment of replication success. The original study
  from \citet{Pyc2010} has effect estimate $\hat \theta_o= 0.4$ on Fisher's $z$
  scale (95\% CI from $0.05$ to $0.74$) and one-sided $p$-value $p_o = 0.011$.
  The left part of the figure illustrates the reverse-Bayes derivation of the
  sufficiently sceptical prior based on the original study result and the
  posterior with lower credible limit fixed at zero. The comparison of the
  sufficiently sceptical prior with the replication study result
  ($\hat \theta_r = 0.15$, 95\% CI from 0.04 to 0.26, $p_r = 0.004$) in the
  right part of the figure is used to assess potential prior-data conflict.}
  \label{fig2:fig1}
\end{center}
\end{figure}
\end{center}

Hereinafter we focus on the one-sided assessment of replication success to
ensure that replication success can only occur if the original and replication
effect estimates go in the same direction. Figure \ref{fig2:fig1} illustrates the
\citet{Held2020} approach based on a replication study from the \textit{Social
  Sciences Replication Project} \citep{Camerer2018}: the significant original
finding by \citet{Pyc2010} at one-sided level $\alpha=0.025$ is challenged with
a sceptical prior, sufficiently concentrated around zero to make the original
study result no longer convincing \citep{Matthews2001a,Matthews2001b}.
Replication success is then defined as conflict between the sceptical prior and
the result from the replication study in order to disprove the sceptic. Conflict
is quantified by a prior-predictive tail probability $p_{\mbox{\scriptsize
    Box}}$ \citep{Box1980} where a small value $p_{\mbox{\scriptsize
    Box}} \leq \alpha$ defines replication success. In Figure \ref{fig2:fig1} the
original finding is only borderline significant, so the sufficiently sceptical
prior is fairly wide. Furthermore, there is substantial shrinkage
($d = 0.15/0.4 = 0.38$) of the replication effect estimate and therefore hardly
any conflict with the sufficiently sceptical prior (one-sided
$p_{\mbox{\scriptsize Box}}=0.31$). We are thus not able to declare replication
success at level $2.5$\%.

The actual value of $p_{\mbox{\scriptsize Box}}$ is difficult to interpret as it
depends on the level $\alpha$ and does not even exist if the original $p$-value
$p_o$ exceeds $\alpha$. However, \citet{Held2020} showed that if both
$\sign(z_o)=\sign(z_r)$ and
\begin{equation}\label{eq2:extrinsic.p}
\left({z_o^2}/{z_{{\alpha_S}}^2}-1 \right) \left({z_r^2}/{z_{{\alpha_S}}^2}
- 1 \right) \geq c
\end{equation}
hold, replication success at level ${\alpha_S}$ is achieved, where
$z_{{\alpha_S}} = \Phi^{-1}(1-{\alpha_S})$. The requirement
\eqref{eq2:extrinsic.p} can be assessed for any value of
${{\alpha_S}} > \max\{p_o, p_r\}$ and of particular interest is the smallest
possible value of $\alpha_S$ where \eqref{eq2:extrinsic.p} holds, the so-called
\textit{sceptical $p$-value} $p_S$. We are thus interested in the value $z_S^2$
that fulfills
\begin{equation}\label{eq2:extrinsic.p2}
\left({z_o^2}/{z_{S}^2}-1 \right) \left({z_r^2}/{z_{S}^2}
- 1 \right) = c.
\end{equation}
There is a unique solution of \eqref{eq2:extrinsic.p2} which defines the
one-sided {sceptical $p$-value} $p_S= 1-\Phi\left({z_S}\right)$ where
$z_{S} \coloneqq + \surd z_{S}^2$, provided $\sign(z_o)=\sign(z_r)$ holds.
Replication success at level $\alpha_S$ is then achieved if $p_S \leq \alpha_S$.
In the introductory example based on the original study by \citet{Pyc2010}, the
sceptical $p$-value turns out to be $p_S=0.11$.

The sceptical $p$-value has a number of interesting properties, see
\citet[Section 3.1]{Held2020} for details. In particular,
$p_S > \max\{p_o, p_r\}$ always holds with $p_S \downarrow \max\{p_o, p_r\}$ for
$c \downarrow 0$. Furthermore, if the $p$-values $p_o$ and $p_r$ are fixed, the
sceptical $p$-value $p_S$ increases with decreasing relative effect size $d$.
The first property ensures that both the original and the replication study have
to be sufficiently convincing on their own to achieve replication success. The
second property guarantees that shrinkage of the replication effect estimate is
penalized.

The level for replication success $\alpha_S$ has to be distinguished from the
significance level $\alpha$ associated with the ordinary $p$-value.
\citet{Held2020} has used the \textit{nominal level} for replication success
($\alpha_S=\alpha$) for convenience, but in the following we will propose a
recalibration of the procedure along with a new value for $\alpha_S$, the
\textit{golden level} (Section~\ref{sec2:goldenthresh}). The derivation is based
on a property of the required relative effect size for replication success, if
the relative sample size is very large (Section~\ref{sec2:res}). In a nutshell,
the golden level ensures that for original studies which were only borderline
significant ($p_o=\alpha$), replication success is only possible if the
replication
effect estimate is larger than the original one ($d > 1$).\\

\subsection{Relative effect size}\label{sec2:res}
Without loss of generality we now assume that $\hat \theta_o > 0$ and that
$p_o < {\alpha_S}$ has been observed in the original study, otherwise it would
be impossible to achieve replication success at level $\alpha_S$ because $p_S$
is always larger than $p_o$. The condition \eqref{eq2:extrinsic.p} for
replication success can then be re-written as
\begin{equation}\label{eq2:eq.z}
  z_r \geq   z_{{\alpha_S}} \sqrt{1+c/(K-1)} \eqqcolon  \zrmin,
\end{equation}
where $K=z_o^2/z_{{\alpha_S}}^2>1$. The right hand-side of \eqref{eq2:eq.z} is
the minimum replication $z$-value $\zrmin$ required to achieve replication
success. Note that $\zrmin$ increases with increasing $c$, so increasing the
replication sample size leads to a more stringent success requirement $z_r$ and
the corresponding replication $p$-value $p_r$.

Equation \eqref{eq2:eq.z} can be further transformed to a
condition on the relative effect size \eqref{eq2:d}:
\begin{equation}\label{eq2:res}
  d \geq  \frac{\sqrt{1+c/(K-1)}}{ \sqrt{c K}} \eqqcolon \dmin .
\end{equation}
To achieve replication success, the relative effect size must be at least as
large as the right hand-side of \eqref{eq2:res}, the minimum relative effect size
$\dmin$, a function of $K$ and the relative sample size $c$. If the relative
sample size becomes very large, \ie $c \rightarrow \infty$, we have
$\dmin \downarrow \dinfty$ where
\begin{equation}\label{eq2:smallestpossible}
  \dinfty = 1/\sqrt{K (K-1)}
\end{equation}
is the \textit{limiting relative effect size}. This shows that the minimum
relative effect size $\dmin$ in \eqref{eq2:res} does not go to zero for
increasing $c$, so replication success cannot be achieved if the relative effect
size $d$ is smaller or equal to $\dinfty$, no matter how large the replication
study is. In contrast, the corresponding criterion \eqref{eq2:dSig} of the
two-trials rule can be achieved for any positive relative effect size,
regardless of how small, provided the replication sample size is sufficiently
large.

\subsection{The golden level}\label{sec2:goldenthresh}
Significance of both the original and the replication study at level $\alpha$ is
a necessary but not sufficient requirement for replication success at the
nominal level ($\alpha_S = \alpha$). The nominal level may therefore be too
stringent. It is more reasonable to calibrate the procedure in such a way that
to establish replication success, original and replication study do not both
necessarily need to be significant at level $\alpha$, provided that the
replication effect estimate does not shrink compared to the original one. We
therefore choose a level $\alpha_S$ such that a borderline significant original
study ($p_o = \alpha$) cannot lead to replication success if there is shrinkage
$s > 0$ of the replication effect estimate. Mathematically, this translates to
setting $\dinfty = 1$ and $K = z_\alpha^2/z_{\alpha_S}^2$
in~\eqref{eq2:smallestpossible} and leads to the quadratic equation $K(K-1) = 1$
with solution $K = \varphi$ where $\varphi = (\sqrt{5}+1)/2 \approx 1.62$ is
known as the golden ratio. Solving for $z_{\alpha_S}$ gives
$z_{\alpha_S} = z_\alpha / \sqrt{\varphi}$ and the corresponding \textit{golden
  level}
\begin{eqnarray}\label{eq2:golden}
\alpha_S & = & 1 -  \Phi(z_\alpha / \sqrt{\varphi} )
\end{eqnarray}
for replication success. This is our recommended default choice to assess
replication success and we will study its properties in the following in more
detail. For $z_\alpha = 1.96$ (one-sided $\alpha = 0.025$), the golden level is
$\alpha_S =0.062$. In the introductory example shown in Figure \ref{fig2:fig1},
the sceptical $p$-value is $p_S=0.11 > 0.062$, so the replication of the
\citet{Pyc2010} study was not successful.

The golden level \eqref{eq2:golden} is derived from~\eqref{eq2:smallestpossible}
with $\dinfty = 1$. However, we may also use a different value for the limiting
relative effect size $\dinfty$, say $\dinfty=0.8$. Then replication success is
only possible for a borderline significant result ($p_o = \alpha$) if there is
less than $1-\dinfty$ (20\% for $\dinfty=0.8$) shrinkage of the replication
effect estimate. This approach is equivalent to a limiting relative effect size
of 1 if the original $p$-value $p_o$ is equal to a different level $\alpha'$,
which can be derived as follows: First, solving \eqref{eq2:smallestpossible} for
$\dinfty > 0$ gives
$K = z_\alpha^2/z_{\alpha_S}^2 = 1/2+\sqrt{1/4 + 1/\dinfty^2}$. The new level
$\alpha'$ fulfills $\varphi=z_{\alpha'}^2/z_{\alpha_S}^2$, so
$z_\alpha^2/K=z_{\alpha'}^2/\varphi$ and therefore
\begin{equation}\label{eq2:alphaPrime}
\alpha' = 1-\Phi\left(z_{\alpha} \, \sqrt{\varphi / K} \right).
\end{equation}
For example, for $\alpha = 0.025$ and $\dinfty = 0.8$ we obtain $\alpha'=0.033$.

\subsection{Recalibration of the sceptical \textit{p}-value}\label{sec2:recalib}

The condition $p_S \leq \alpha_S$ for replication success at the golden level is
equivalent to $z_S \geq z_\alpha / \sqrt{\varphi}$, \ie
$z_S \sqrt{\varphi} \geq z_\alpha$. In practice it may be preferable to
recalibrate the sceptical $p$-value $p_{S} = 1 - \Phi(z_S)$ to
$ \tilde{p}_S = 1 - \Phi(z_S \sqrt{\varphi})$, which then needs to be compared
to $\alpha$ (rather than $\alpha_S$) to assess replication success and can thus
be interpreted on the same scale as an ordinary $p$-value. For example, the
recalibrated sceptical $p$-value for the replication of \citet{Pyc2010} turns
out to be $\tilde{p}_S=0.061$ and does not lead to replication success at any
level $\alpha<0.061$, including the standard $0.025$ level.

\subsection{Comparison with the two-trials rule}\label{sec2:2TR}
A useful benchmark for comparison is the two-trials rule in drug development
\citep[Section 9.4]{Kay2015}, which requires ``at least two adequate and
well-controlled studies, each convincing on its own, to establish
effectiveness'' \citep[p.~3]{FDA1998}. This is usually achieved by independently
replicating the result of a first study in a second study, both significant at
one-sided level $\alpha=0.025$. {It is worth noting that in practice the two
  trials are often run in parallel} {\citep{Senn2008}}{, so do not exactly
  resemble the replication setting.}

The main difference between the replication success and the two-trials
rule approach concerns how shrinkage of the replication effect
estimate is handled. Figure~\ref{fig2:fig2} illustrates that
shrinkage is penalized in the assessment of replication success, \ie
the original $p$-value needs to be quite small to achieve replication
success for a relative effect size $d<1$. In contrast, significance of
the replication study can be achieved even if there is substantial
shrinkage, provided the replication sample size is large enough.

\begin{figure}[!h]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\maxwidth]{images/paper2/fig2-1}

}

\end{knitrout}
\caption{Comparison of replication success at the golden level
  ($p_S \leq \alpha_S = 0.062$) and the two-trials rule ($p_o \leq 0.025$ and
  $p_r \leq 0.025$). The dotted areas indicate that success is impossible for
  original $p$-value $p_o$ and relative effect size $d$. In the white areas
  success is possible and depends on the relative sample size $c$ as indicated
  by the grey lines. The dashed black line in the left plot indicates the
  limiting relative effect size $\dinfty$. }
\label{fig2:fig2}
\end{figure}
